\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{comment}
\usepackage{microtype}
\usepackage{bm}
\usepackage{multicol}

\usepackage{mathtools}
\DeclarePairedDelimiter\bra{\langle}{\rvert}
\DeclarePairedDelimiter\ket{\lvert}{\rangle}
\DeclarePairedDelimiterX\braket[2]{\langle}{\rangle}{#1\delimsize\vert\mathopen{}#2}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\DeclarePairedDelimiterX\expval[3]{\langle}{\rangle}%
{#1\delimsize\vert\mathopen{}#2\delimsize\vert\mathopen{}#3}
\newcommand\HF{\ensuremath{\mathrm{HF}}}

\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{ % this is just my personal choice, feel free to change things
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black},
}

\usepackage{simpler-wick}

\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}

\renewcommand\thesection{Part \alph{section})}  % chktex 9 % chktex 10 % tex-fmt: skip
\renewcommand\thesubsection{Solution} % tex-fmt: skip
\renewcommand\thesubsubsection{\arabic{subsection}.\arabic{subsubsection})} % chktex 9 % chktex 10 % tex-fmt: skip

\title{
    First midterm FYS4480\\
    Quantum mechanics for many-particle systems
}
\author{August Femtehjell \& Oskar Idland}
\date{October 2024}

\begin{document}

\maketitle

\section*{Introduction}

In this midterm we will develop two simple models for studying the helium atom (with two electrons) and the beryllium atom with four electrons.

After having introduced the  Born-Oppenheimer approximation which effectively freezes out the nucleonic degrees of freedom, the Hamiltonian for \(N\) electrons takes the following form
\begin{equation*}
    \hat{H} = \sum_{i=1}^{N} t(x_i) - \sum_{i=1}^{N} k\frac{Ze^2}{r_i} + \sum_{i<j}^{N} \frac{ke^2}{r_{ij}},
\end{equation*}
with \(k=1.44\) eVnm.
Throughout this work we will use atomic units, this means that \(\hbar = c = e = m_e = 1\).
The constant \(k\) becomes also equal 1.
The resulting energies have to be multiplied by \(2 \times 13.6\) eV in order to obtain energies in eletronvolts.

We can rewrite our Hamiltonians as
\begin{equation}
    \hat{H} = \hat{H_0} + \hat{H_I}
    = \sum_{i=1}^{N}\hat{h}_0(x_i) + \sum_{i<j}^{N}\frac{1}{r_{ij}},
    \label{eq:H1H2}
\end{equation}
where  we have defined \(r_{ij} = |\boldsymbol{r}_i - \boldsymbol{r}_j|\) and \(\hat{h}_0(x_i) =  \hat{t}(x_i) - \frac{Z}{r_i}\).

The variable \(x\) contains both the spatial coordinates and the spin values.
The first term of Eq.~\eqref{eq:H1H2}, \(H_0\), is the sum of the \(N\) \emph{one-body} Hamiltonians \(\hat{h}_0\).
Each individual Hamiltonian \(\hat{h}_0\) contains the kinetic energy operator of an electron and its potential energy due to the attraction of the nucleus.
The second term, \(H_I\), is the sum of the \(N(N-1)/2\) two-body interactions between each pair of electrons.
Note that the double sum carries a restriction \(i<j\).

As basis functions for our calculations we will use hydrogen-like single-particle functions.
This means the onebody operator is diagonal in this basis for states \(i,j\) with quantum numbers
\(n,l,m_l,s,m_s\) with energies
\begin{equation}
    \expval{i}{\hat{h}_0}{j} = -\frac{Z^2}{2n^2}\delta_{ij}.
    \label{eq:onebody}
\end{equation}
The quantum number \(n\) refers to the number of nodes of the wave function.
Observe that this expectation value is independent of spin.

We will in all calculations here restrict ourselves to only so-called \(s\)-waves, that is the orbital momentum \(l\) is zero.
We will also limit the quantum number \(n\) to \(n \le 3\).
It means that every \(ns\) state can accommodate two electrons due to the spin degeneracy.

In the calculations you will need the Coulomb interaction with matrix elements involving single-particle wave functions with \(l = 0\) only, the
so-called \(s\)-waves.
We need only the radial part since the spherical harmonics for the \(s\)-waves are rather simple.
We omit single-particle states with \(l > 0\).
The actual integrals we need, are tabulated at the end.
Our radial wave functions are
\begin{equation*}
    R_{n0}(r) = \left( \frac{2Z}{n} \right)^{3/2} \sqrt{\frac{(n-1)!}{2n\times n!}} L_{n-1}^1 \left( \frac{2Zr}{n} \right) \exp{\left( -\frac{Zr}{n} \right)},
\end{equation*}
where \(L_{n-1}^1(r)\) are the so-called Laguerre polynomials.
These wave functions can then be used to compute the direct part of the Coulomb interaction
\begin{equation*}
    \expval*{\alpha\beta}{V}{\gamma\delta} = \int r_1^2 dr_1 \int r_2^2 dr_2 R_{n_{\alpha} 0}^*(r_1) R_{n_{\beta} 0}^*(r_2) \frac{1}{r_{12}}R_{n_{\gamma} 0}(r_1) R_{n_{\delta} 0}(r_2).
\end{equation*}

Observe that this is only the radial integral and that the labels \( \alpha,\beta,\gamma,\delta \) refer only to the quantum numbers \(n,l,m_l\), with \(m_l\) the projection of the orbital momentum \(l\).
A similar expression can be found for the exchange part.
Since we have restricted ourselves to only \(s\)-waves, these integrals are straightforward but tedious to calculate.
As an addendum to this midterm we list all closed-form expressions for the relevant matrix elements.
Note well that these matrix elements do not include spin.
When setting up the final antisymmetrized matrix elements you need to consider the spin degrees of freedom as well.
Please pay in particular attention to the exchange part and the pertinent spin values of the single-particle states.

We will also, for both helium and beryllium assume that the many-particle states we construct have always the same total spin projection \(M_S = 0\).
This means that if we excite one or two particles from the ground state, the spins of the various single-particle states should always sum up to zero.

\section{Setting up the basis}
\input{sections/a.tex}

\section{Second quantized Hamiltonian}
% \input{sections/b.tex}
\input{sections/b_new.tex}
\section{Limiting ourselves to one-particle-one excitations}
\input{sections/c_new.tex}

\section{Moving to the Beryllium atom}
\input{sections/d.tex}

\section{Hartree-Fock}
\subsection*{Preamble}
With a given energy functional, we can perform at least two types of variational strategies.
These are:
\begin{enumerate}
    \item Vary the Slater determinant by changing the spatial part of the single-particle wave functions themselves, or

    \item Expand the single-particle functions in a known basis  and vary the coefficients, that is, the new function single-particle wave function $\ket{p}$ is written as a linear expansion in terms of a fixed basis $\phi$ (harmonic oscillator, Laguerre polynomials etc)
        \begin{equation*}
            \psi_p  = \sum_{\lambda} C_{p\lambda}\phi_{\lambda}
        \end{equation*}
\end{enumerate}
Both cases lead to a new Slater determinant which is related to the previous via a unitary transformation.
Below we will set up the Hartree-Fock equations using the second option.
We assume that our basis is still formed by the hydrogen-like wave functions.
We consider a Slater determinant built up of single-particle orbitals $\phi_{\lambda}$ where the indices $\lambda$ refer to specific single-particle states.
As an example, you could think of the ground state ansatz for the beryllium atom.

The unitary transformation
\begin{equation*}
    \psi_p  = \sum_{\lambda} C_{p\lambda}\phi_{\lambda},
\end{equation*}
brings us into the new basis $\psi$.
The new basis is orthonormal and $C$ is a unitary matrix.

\subsection*{Problem definition}
\input{sections/e.tex}

\end{document} % chktex 17
