The Hartree-Fock equations with a variation of the coefficients $C_{p\alpha}$ lead to an eigenvalue problem whose eigenvectors are the coefficients $C_{p\alpha}$ and eigenvalues are the new single-particle energies.
Use the single-particle orbits $1s{-}3s$ and set up the Hartree-Fock matrix for both the helium atom and the beryllium atom.
Find after the first diagonalization the new single-particle energies and the new ground state energy.
Compare these results with those you obtained under the minimization of the ground states as functions of $Z$ and the full diagonalization.
When setting up the Hartree-Fock matrix in the first iteration, our guess for the coefficients $C_{p\beta}$ etc.\ is $C_{p\beta}=1$ for $p=\beta$ and zero else.

\subsection{}
We set up the Hartree-Fock matrices with the code in \verb|hartree_fock.py|, for both the helium and beryllium atoms.
After the first diagonalization, we find the new single-particle energies and the new ground state energy, listed in \autoref{tab:energies}.
The untruncated values can be found in \verb|hartree_fock.txt|.

\begin{table}[h!]
    \caption{Single-particle energies and new ground state energies for Helium and Beryllium atoms, after one Hartree-Fock iteration.\label{tab:energies}}
    \centering
    \small
    \begin{tabular}{lcr}
        \toprule\toprule
        Atom & Single-Particle Energies & New Energy \\ \midrule
        Helium & $
        \begin{bmatrix}
            -0.7832 & -0.7832
        \end{bmatrix}
        $ & $-2.8291$ \\ \midrule
        Beryllium & $
        \begin{bmatrix}
            -3.9506 & -3.9506 & -0.1040 & -0.1040
        \end{bmatrix}
        $ & $-14.4998$ \\ \bottomrule
    \end{tabular}
\end{table}

For the helium atom, we find that the new ground state energy is $-2.8291$ atomic units, which is an improvement from the value $-2.75$ atomic units we found when minimizing with respect to $Z$, but slightly worse than the result $-2.8386$ atomic units we found when fully diagonalizing the Hamiltonian.

For the beryllium atom, the new ground state energy of $-14.4998$ atomic units is a large improvement from the estimate of $-13.7159$ we got when minimizing with respect to $Z$, and is still better than the fully diagonalized result of $-14.3621$ atomic units.
